\section{Conclusion}
\label{sec:conclusion}

We introduced CLP-CSGM Ridge, a conditional generative compressed sensing method
for petrophysical property estimation from dense logs and sparse core
measurements. The method learns a target-property decoder, predicts a
log-conditioned latent prior with ridge regression, and refines the latent code
at test time by enforcing sparse measurement consistency.

In cross-well low-data Vc prediction, CLP-CSGM Ridge outperformed the strongest
direct \([u,b]\) autoencoder baseline in all tested training-size and
measurement-ratio cells. In the real-well F03 GR-only porosity case, it
achieved the best average RMSE, with its clearest advantage in the sparsest
measurement regimes. These results indicate that conditional latent generative
priors are a promising alternative to fixed-basis sparse residual recovery for
petrophysical sparse-measurement assimilation.

The ablation results clarify why the method works in the studied regimes. The
prior-only decoder \(G(h(u))\) is weaker than the full method, showing that
sparse target-well measurements add information beyond log-conditioned latent
regression. Measurement-only CSGM is also weaker, showing that sparse
observations alone do not provide a stable latent estimate in these low-data
settings. Finally, the Ridge prior is consistently stronger than the MLP prior,
supporting the use of a low-capacity conditional latent prior when training
windows are scarce.

Future work should evaluate the method on additional fields, compare richer
conditional priors, investigate uncertainty quantification in latent space, and
study measurement designs beyond coordinate subsampling.
