\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,calc}

\graphicspath{{figures/}}

\newtheorem{assumption}{Assumption}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\title{Conditional Latent-Prior Generative Compressed Sensing for Petrophysical Property Estimation from Sparse Core Measurements}
\author{Author names omitted for review}
\date{}

\begin{document}

\maketitle

\begin{abstract}
Dense well logs provide continuous indirect observations of subsurface
properties, whereas core measurements are sparse, expensive, and usually
available only at selected depths. Direct supervised models can exploit log
features, but they often use sparse target-well measurements only as additional
inputs. Classical compressed sensing provides measurement consistency, but its
fixed-basis sparsity assumption can be too rigid for cross-well petrophysical
prediction. We propose Conditional Latent-Prior Compressed Sensing with
Generative Models (CLP-CSGM), which learns a generative range of
target-property windows and uses well logs to predict a latent prior. Sparse
core measurements then refine the reconstruction through a test-time latent
inverse problem. In low-data cross-well clay-volume prediction, CLP-CSGM Ridge
improves over a direct autoencoder baseline in all tested training-size and
measurement-ratio cells, reducing mean RMSE from 0.01905 to 0.01397. In a
real-well F03-4 GR-only porosity stress test, CLP-CSGM Ridge attains the best
average RMSE, with strongest gains in the sparsest measurement regimes.
Ablations confirm that both the conditional prior and the sparse-measurement
term are necessary, and that Ridge is more stable than an MLP latent prior in
the tested low-data grids.
\end{abstract}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_method}
\input{sections/04_data_and_eda}
\input{sections/05_experiments}
\input{sections/06_results}
\input{sections/07_discussion}
\input{sections/08_conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
